The Inner Mass-Ratio Distribution of Intermediate-Mass Stars

### 2016-03-08 paper/ms.tex
Updated sections 4-5

In addition to the spectroscopic follow-up, we obtained Gemini/NIRI adaptive optics imaging data for 18 of the northern companions, and were able to resolve 7 of them. We show the separation, position angle, and magnitude difference measurements in Table \ref{tab:imaging_obs}, and display the images in Figure \ref{fig:images}. We also derive the projected separation in AU and the companion mass from the images. We calculate the separation from the measured angular separation and the Hipparchos parallax \citep{Hipparchos}. We calculate the companion mass and uncertainties from 1000 samples of the magnitude difference measurement and the primary star mass, temperature, age, and radius (see Section \ref{sec:sp}). For each sample, we use a grid of Kurucz stellar model spectra \citep{Castelli2003}and the pysynphot code\footnote{pysynphot is a python package to perform synthetic photometry, and is available at this url: \url{https://pypi.python.org/pypi/pysynphot}} to determine the companion temperature needed to replicate the observed magnitude difference. We estimate the companion radius by interpolating solar metallicity Dartmouth isochrones [CITE] from the companion temperature and system age sample. We convert the best temperature to a companion mass using the same isochrone grid. The masses derived from the imaging data have very large uncertainties because the primary star property estimates that they depend on are very uncertain. The imaging masses agree with the spectroscopically-derived masses in Table \ref{tab:companions}, with the exception of HIP 115115 which has a much higher mass from the imaging data than the spectroscopic data. The spectroscopic masses for all stars are more reliable, since they are less model-dependent.

@@

\section{Primary Star Parameters}
\label{sec:sp}

@@

About half of our detections have robust mass and age estimates for the primary stars from Str\"omgren uvby$\beta$ photometry \citep{David2015}. For those that do not, we estimate the mass and age of the system from our spectra. We first cross-correlate the data against a grid of model spectra spanning

\begin{itemize}
\item $7000\ K < T_\mathrm{eff} < 30000\ K$ in steps of $1000\ K$
\item $3.0 < \log{g} < 4.5$ in steps of $0.5$ dex
\item $75 < v\sin{i} < 300\ \mathrm{km s}^{-1}$ in steps of $25\ \mathrm{km s}^{-1}$
\end{itemize}
Since the Phoenix model library does not extend to temperatures $T > 12000\ K$, we used Kurucz model spectra \citep{Castelli2003} for this grid.

For the optical data, we use the blue \'echelle orders ($\lambda < 5550 \AA$). We ignore the strong hydrogren Balmer lines in the spectrum because they span several \'echelle orders and make continuum normalization very difficult, potentially biasing the result. There are sufficient metal lines in the optical spectra that the resulting CCF always has a very strong peak at the radial velocity of the primary star. The near-infrared IGRINS spectra have very few strong metal lines; we use the subset from $1.51 - 1.73 \micron$ that is dominated by hydrogren Brackett lines for these spectra.  We then estimate the temperature and surface gravity of the primary stars from the CCF with the largest peak. We adopt the following errors on the temperature and surface gravity:

\begin{align}
 \sigma_T &= \begin{cases}
      \hfill 500\ K \hfill & T < 10000\ K \\
      \hfill 1000\ K \hfill & T >= 10000\ K \\
     \end{cases} \\
 \sigma_{\log{g}} &= 0.25
\end{align}
The IGRINS parameters are less reliable because they rely almost solely on the hydrogren Brackett lines that span an entire \'echelle order, so we double the uncertainty on the IGRINS-derived temperature, surface gravity, and metallicity. Additionally, we throw out the IGRINS parameters if the star was also observed by one of the optical instruments in our survey. For stars observed multiple times, we use the average parameters and reduce the uncertainties accordingly.

@@

We show the temperature, surface gravity, mass, and age estimates for most of our sample stars Table \ref{tab:sample}. We do not give parameters for the few stars that show strong discrepancies with the spectral-type estimate, most of which are early B-stars that have temperatures higher than the maximum grid temperature of $30000\ K$. The MCMC samples generated by the isochrones code are available at the following url: \url{https://zenodo.org/record/46482}.

@@

Next, we cross correlated all of the synthetic observations against the Phoenix model template that was used to construct them. We counted the companion as detected if the highest point in the resulting CCF was found at the correct radial velocity, and if the peak had a significance of $>5\sigma$, where $\sigma$ is the standard deviation of the CCF for points more than $100\ \mathrm{km\ s}^{-1}$ away from the peak. We then combined all of the radial velocity trials for each grid point to estimate a probability of detection at that grid point:

@@

Finally, we interpolated between the grid points using a linear radial basis function interpolator (Figure \ref{fig:detrate_2d}). In order to extrapolate from our grids to estimate the detection rate at high temperature and low $v\sin{i}$ and at low temperature and high $v\sin{i}$, we made the following assumptions about the shape of the two-dimensional detection rate surface: First, we assumed that if all companions are detected at temperature $T=6500$ K and rotation speed $v\sin{i}$, then all companions with the same $v\sin{i}$ and larger temperature will also be detected (lower right points in Figure \ref{fig:detrate_2d}). Likewise, we assume that if no companions are detected at temperature $T$ and rotation speed $v\sin{i} = 50\ \mathrm{km\ s}^{-1}$, then no companions will be detected at the same temperature and faster rotation speeds (upper left points in Figure \ref{fig:detrate_2d}). Figure \ref{fig:detrate_2d} shows a clear diagonal dividing line between hot, slow rotators that are always detected and cool, fast rotators that never are. Additionally the figure shows that very fast rotators are never detected, regardless of their temperature, because the signal is completely removed when we high-pass filter the data (see Section \ref{sec:companions}).

@@

\begin{equation}
Q(T) = \sum_k Q(T, v_k) v_k 
\end{equation} 
where $Q(T, v\sin{i})$ is the surface plotted in Figure \ref{fig:detrate_2d} and $v_k$ are the samples from the distribution of $v\sin{i}$. For $T < 6000$ K, we sample $v\sin{i}$ using the gyrochronology relation given in \citet{Barnes2010b}:

@@

\begin{itemize}
\item $k_C = 0.646$ day/Myr
\item $k_I = 452$ Myr/day
\end{itemize}
We estimate the convective timescale ($\tau$) by interpolating Table 1 of \citet{Barnes2010a}. We then randomly sample the system age $t$ from its probability distribution function (see Section \ref{sec:sp} and Table \ref{tab:sample}). Young stars have rotation periods in the range of 0.2 to 10 days \citep{Bouvier2014}, so we randomly choose an initial rotation period $P_0$ from a log-uniform distribution in this range for each age sample. Equation \ref{eqn:gyro} then gives a current rotation period for each sample, which we convert to an equatorial velocity with the stellar radius $R$. We estimate the stellar radius by interpolating Dartmouth pre main sequence isochrones [CITE] using the companion temperature and system age. We finally convert to projected velocity $v\sin{i}$ by randomly sampling a uniform distribution for the inclination $\sin{i}$.

@@

We test to determine if the completeness is reasonable at large mass ratios by comparing to known binary systems. We detect 15 of the 25 stars in our sample with a hot ($T > 7000\ K$) companion. The completeness function for hot, roughly equal-mass companions suggest the probability of detection if $\sim 80\%$, which is incompatible with our low detection rate (p=0.01). The discrepancy may be due to an underestimate of the typical rotation rates for hot stars, used when marginalizing out the dependence on $v\sin{i}$. Additionally, rapidly rotating companions, especially when they have a similar temperature to the primary, are more difficult to detect if they have a small radial velocity offset from the primary star. While the injection and recovery experiments do sample velocity space to account for this, they may be over-sampling companions with very large velocity offsets and producing anomalously high detection rates. We account for the discrepancy by introducing a scaling factor: we multiply the estimated detection rate for all companions with $T > 7000\ K$ by $f=0.8$.

@@

We are now finally in a position to estimate the mass-ratio distribution for our sample. We estimate the mass for each detected companion star by sampling the temperatures given in Table \ref{tab:companions} as a gaussian, and converting each temperature sample into a mass sample. We do the conversion to mass both by interpolating Table 5 from \citet{Pecaut2013}, and by interpolating from temperature and system age (see Section \ref{sec:sp}) to mass with Dartmouth isochrones [CITE]. Both methods give similar results in most cases, although the isochrone masses are significantly higher for very young systems, so we adopt the isochronal values in the analysis. We sample the mass ratio of the system by dividing the companion mass samples by samples of the primary mass (Section \ref{sec:sp}). We denote the $n$th mass ratio sample for the $k$th star as $q_k^{(n)}$, and denote the number of these samples as $N_k$.

@@

In this work, we described a binary survey of 341 bright A- and B-type stars. We used the direct spectral detection method \citep{Gullikson2016} to find the spectral lines of 64 companions with temperatures ranging from $3600 - 16000\ K$.  We used the cross-correlation functions to estimate the temperature, surface gravity, and metallicity of most of our sample stars, and converted to mass and age by interpolating Padova stellar evolutionary tracks \citep{Bressan2012}. Likewise, we convert the companion temperature measurements to mass by using solar metallicity Dartmough evolutionary tracks [CITE].

>>>>

In addition to the spectroscopic follow-up, we obtained Gemini/NIRI adaptive optics imaging data for 18 of the northern companions, and were able to resolve 7 of them. We show the separation, position angle, and magnitude difference measurements in Table \ref{tab:imaging_obs}, and display the images in Figure \ref{fig:images}. We also derive the projected separation in AU and the companion mass from the images. We calculate the separation from the measured angular separation and the Hipparchos parallax \citep{Hipparchos}. We calculate the companion mass and uncertainties from 1000 samples of the magnitude difference measurement and the primary star mass, temperature, age, and radius (see Section \ref{sec:sp}). For each sample, we use a grid of Kurucz stellar model spectra \citep{Castelli2003}and the pysynphot code\footnote{pysynphot is a python package to perform synthetic photometry, and is available at this url: \url{https://pypi.python.org/pypi/pysynphot}} to determine the companion temperature needed to replicate the observed magnitude difference. We estimate the companion radius by interpolating solar metallicity Dartmouth isochrones \citep{Dotter2008} from the companion temperature and system age sample. We convert the best temperature to a companion mass using the same isochrone grid. The masses derived from the imaging data have very large uncertainties because the primary star property estimates that they depend on are very uncertain. The imaging masses agree with the spectroscopically-derived masses in Table \ref{tab:companions}, with the exception of HIP 115115 which has a much higher mass from the imaging data than the spectroscopic data. The spectroscopic masses for all stars are more reliable, since they are less model-dependent.

@@

\section{Sample Star Parameters}
\label{sec:sp}

@@

About half of our sample stars have robust mass and age estimates from Str\"omgren uvby$\beta$ photometry \citep{David2015}. For those that do not, we estimate the mass and age of the system from our spectra. We first cross-correlate the data against a grid of solar metallicity Kurucz model spectra \citep{Castelli2003} spanning

\begin{itemize}
\item $7000\ K < T_\mathrm{eff} < 30000\ K$ in steps of $1000\ K$
\item $3.0 < \log{g} < 4.5$ in steps of $0.5$ dex
\item $75 < v\sin{i} < 300\ \mathrm{km s}^{-1}$ in steps of $25\ \mathrm{km s}^{-1}$
\end{itemize}

For the optical data, we use the blue \'echelle orders ($\lambda < 5550 \AA$). We ignore the strong hydrogren Balmer lines in the spectrum because they span several \'echelle orders and make continuum normalization very difficult, potentially biasing the result. There are sufficient metal lines in the optical spectra that the resulting CCF always has a very strong peak at the radial velocity of the primary star. The near-infrared IGRINS spectra have very few strong metal lines; we use the subset from $1.51 - 1.73 \micron$ that is dominated by hydrogren Brackett lines for these spectra. Similar to the companion search, we estimate the temperature and surface gravity of the stars from the CCF with the largest peak. We adopt the following errors on the temperature and surface gravity:

\begin{align}
 \sigma_T &= \begin{cases}
      \hfill 500\ K \hfill & T < 10000\ K \\
      \hfill 1000\ K \hfill & T >= 10000\ K \\
     \end{cases} \\
 \sigma_{\log{g}} &= 0.25
\end{align}
The IGRINS parameters are less reliable because they rely almost solely on the hydrogren Brackett lines that span an entire \'echelle order, so we double the uncertainty on the IGRINS-derived temperature and surface gravity. Additionally, we throw out the IGRINS parameters if the star was also observed by one of the optical instruments in our survey. For stars observed multiple times, we use the average parameters and reduce the uncertainties accordingly.

@@

We show the temperature, surface gravity, mass, and age estimates for most of our sample stars Table \ref{tab:sample}. We do not give parameters for the few stars that show strong discrepancies with the spectral-type estimate, most of which are early B-stars that have temperatures higher than the maximum grid temperature of $30000\ K$.

@@

Next, we cross correlated all of the synthetic observations against the Phoenix model template that was used to construct them. We counted the companion as detected if the highest point in the resulting CCF was found at the correct radial velocity, and if the peak had a significance of $>5\sigma$, where $\sigma$ is the standard deviation of the CCF for points more than $100\ \mathrm{km\ s}^{-1}$ away from the peak. We combined all of the radial velocity trials for each grid point to estimate a probability of detection at that grid point:

@@

Finally, we interpolated between the grid points using a linear radial basis function interpolator (Figure \ref{fig:detrate_2d}). In order to extrapolate from our grids to estimate the detection rate at high temperature and low $v\sin{i}$ and at low temperature and high $v\sin{i}$, we made the following assumptions about the shape of the two-dimensional detection rate surface: First, we assumed that if all companions are detected at temperature $T=6500$ K and rotation speed $v\sin{i}$, then all companions with the same $v\sin{i}$ and larger temperature will also be detected (lower right points in Figure \ref{fig:detrate_2d}). Likewise, we assume that if no companions are detected at temperature $T$ and rotation speed $v\sin{i} = 50\ \mathrm{km\ s}^{-1}$, then no companions will be detected at the same temperature and faster rotation speeds (upper left points in Figure \ref{fig:detrate_2d}). We tested the former assumption with injection and recovery experiments on a small subset of stars, and found that it is valid.

Figure \ref{fig:detrate_2d} shows a clear diagonal dividing line between hot, slow rotators that are always detected and cool, fast rotators that never are. Additionally the figure shows that very fast rotators are never detected, regardless of their temperature, because the signal is completely removed when we unsharp-mask the data (see Section \ref{sec:companions}).

@@

\begin{equation}
Q(T) = \sum_k Q(T, v_k) v_k 
\end{equation} 
where $Q(T, v)$ is the surface plotted in Figure \ref{fig:detrate_2d} and $v_k$ are the samples from the distribution of $v\sin{i}$. For $T < 6000$ K, we sample $v\sin{i}$ using the gyrochronology relation given in \citet{Barnes2010b}:

@@

\begin{itemize}
\item $k_C = 0.646$ day/Myr
\item $k_I = 452$ Myr/day
\end{itemize}
We estimate the convective timescale ($\tau$) by interpolating Table 1 of \citet{Barnes2010a}. We then randomly draw a system age $t$ from its probability distribution function (see Section \ref{sec:sp} and Table \ref{tab:sample}). Young stars have rotation periods in the range of 0.2 to 10 days \citep{Bouvier2014}, so we randomly choose an initial rotation period $P_0$ from a log-uniform distribution in this range for each age sample. Equation \ref{eqn:gyro} then gives a current rotation period for each sample, which we convert to an equatorial velocity with the stellar radius $R$. We estimate $R$ by interpolating Dartmouth pre main sequence isochrones \citep{Dotter2008} using the companion temperature and system age. We finally convert to projected velocity $v\sin{i}$ by randomly sampling a uniform distribution for the inclination $\sin{i}$.

@@

We test to determine if the completeness is reasonable at large mass ratios by comparing to known binary systems. We detect 15 of the 25 stars in our sample with a hot ($T > 7000\ K$) companion. The completeness function for hot, roughly equal-mass companions suggest the probability of detection if $\sim 80\%$, which is incompatible with our low detection rate. The discrepancy may be due to an underestimate of the typical rotation rates for hot stars, which we use when marginalizing out the dependence on $v\sin{i}$. Additionally, rapidly rotating companions, especially when they have a similar temperature to the primary, are more difficult to detect if they have a small radial velocity offset from the primary star. While the injection and recovery experiments do sample velocity space to account for this, they may be over-sampling companions with very large velocity offsets and producing anomalously high detection rates. We account for the discrepancy by introducing a scaling factor: we multiply the estimated detection rate for all companions with $T > 7000\ K$ by $f=0.8$.

@@

We are now finally in a position to estimate the mass-ratio distribution for our sample. We estimate the mass for each detected companion star by sampling the temperatures given in Table \ref{tab:companions} as a gaussian, and converting each temperature sample into a mass sample. We do the conversion to mass both by interpolating Table 5 from \citet{Pecaut2013}, and by interpolating from temperature and system age (see Section \ref{sec:sp}) to mass with Dartmouth isochrones \citep{Dotter2008}. Both methods give similar results in most cases, although the isochrone masses are significantly higher for very young systems, so we adopt the isochronal values in the analysis. We sample the mass ratio of the system by dividing the companion mass samples by samples of the primary mass (Section \ref{sec:sp}). We denote the $n$th mass ratio sample for the $k$th star as $q_k^{(n)}$, and denote the number of these samples as $N_k$.

@@

In this work, we described a binary survey of 341 bright A- and B-type stars. We used the direct spectral detection method \citep{Gullikson2016} to find the spectral lines of 64 companions with temperatures ranging from $3600 - 16000\ K$.  We used the cross-correlation functions to estimate the temperature, surface gravity, and metallicity of most of our sample stars, and converted to mass and age by interpolating Padova stellar evolutionary tracks \citep{Bressan2012}. Likewise, we convert the companion temperature measurements to mass by using solar metallicity Dartmough evolutionary tracks \citep{Dotter2008}.
